# Data Visualization {#sec-viz}

```{r}
#| label: setup-viz
#| include: false 
#| purl: false

# for number learning checks
chap <- 2
lc <- 0

# `r paste0(chap, ".", (lc <- lc + 1))`


knitr::opts_chunk$set(
  tidy = FALSE, 
  out.width = '\\textwidth', 
  fig.height = 4,
  fig.align='center',
  warning = FALSE
)

options(scipen = 99, digits = 3)

# In knitr::kable printing replace all NA's with blanks
options(knitr.kable.NA = '')

# Set random number generator see value for replicable pseudorandomness. Why 76?
# https://www.youtube.com/watch?v=xjJ7FheCkCU
set.seed(76)
```

We begin the development of your data science toolbox with data visualization. By visualizing our data, we gain valuable insights that we couldn't initially see from just looking at the raw data in spreadsheet form.  We will use the `ggplot2` package as it provides an easy way to customize your plots. `ggplot2` is rooted in the data visualization theory known as _The Grammar of Graphics_ [@wilkinson2005].

At the most basic level, graphics/plots/charts (we use these terms interchangeably in this book) provide a nice way for us to get a sense for how quantitative variables compare in terms of their center (where the values tend to be located) and their spread (how they vary around the center).  Graphics should be designed to emphasize the findings and insight you want your audience to understand.  This does however require a balancing act.  On the one hand, you want to highlight as many meaningful relationships and interesting findings as possible; on the other you don't want to include so many as to overwhelm your audience.  

As we will see, plots/graphics also help us to identify patterns and outliers in our data.  We will see that a common extension of these ideas is to compare the *distribution* of one quantitative variable (i.e., what the spread of a variable looks like or how the variable is *distributed* in terms of its values) as we go across the levels of a different categorical variable.

## Packages Needed {.unnumbered}

Let's load all the packages needed for this chapter (this assumes you've already installed them). Read @sec-packages for information on how to install and load R packages.

```{r}
#| message: false

library(nycflights13)
library(ggplot2)
library(dplyr)
```

```{r}
#| message: false
#| warning: false
#| echo: false

# Packages needed internally, but not in text.
library(gapminder)
library(knitr)
library(kableExtra)
library(readr)
```

## The Grammar of Graphics {#sec-grammarofgraphics}

We begin with a discussion of a theoretical framework for data visualization known as "The Grammar of Graphics," which serves as the foundation for the `ggplot2` package. Think of how we construct sentences in English to form sentences by combining different elements, like nouns, verbs, particles, subjects, objects, etc. However, we can't just combine these elements in any arbitrary order; we must do so following a set of rules known as a linguistic grammar. Similarly to a linguistic grammar, "The Grammar of Graphics" define a set of rules for constructing *statistical graphics* by combining different types of *layers*. This grammar was created by Leland Wilkinson [@wilkinson2005] and has been implemented in a variety of data visualization software including R. 

### Components of the Grammar

In short, the grammar tells us that:

> **A statistical graphic is a `mapping` of `data` variables to `aes`thetic attributes of `geom`etric objects.**

Specifically, we can break a graphic into three essential components:

1. `data`: the data set composed of variables that we map.
1. `geom`: the geometric object in question. This refers to the type of object we can observe in a plot. For example: points, lines, and bars.
1. `aes`: aesthetic attributes of the geometric object. For example, x-position, y-position, color, shape, and size.  Each assigned aesthetic attribute can be mapped to a variable in our data set.

You might be wondering why we wrote the terms `data`, `geom`, and `aes` in a computer code type font. We'll see very shortly that we'll specify the elements of the grammar in R using these terms. However, let's first break down the grammar with an example.

### Gapminder data {#sec-gapminder}

```{r}
#| echo: false

gapminder_2007 <- gapminder %>% 
  filter(year == 2007) %>% 
  select(-year) %>% 
  rename(
    Country = country,
    Continent = continent,
    `Life Expectancy` = lifeExp,
    `Population` = pop,
    `GDP per Capita` = gdpPercap
  )
```

In February 2006, a statistician named Hans Rosling gave a TED talk titled ["The best stats you've ever seen"](https://www.ted.com/talks/hans_rosling_shows_the_best_stats_you_ve_ever_seen) where he presented global economic, health, and development data from the website [gapminder.org](http://www.gapminder.org/tools/#_locale_id=en;&chart-type=bubbles). For example, for the `r nrow(gapminder_2007)` countries included from 2007, let's consider only the first 6 countries when listed alphabetically in @tbl-gapminder-2007.


