# Exercise solutions {#sec-exr-sol}

```{r}
library(ISDSdatasets)
```

## Chapter 1 {#sec-ex01-sol}

@exr-ch01-c01 b. Quarto Document

@exr-ch01-c03 a. TRUE

@exr-ch01-c05 b. FALSE

@exr-ch01-c07 b. Data on a flight

@exr-ch01-app1

```{r}
z <- 12*31
add_on <- 12
z + add_on
```

@exr-ch01-adv1

```{r}
head(titanic)
```

The `head()` function shows the first 6 rows of the dataset. Based on this, I expect the `tail()` function to show the last 6 rows of the dataset.

@exr-ch01-adv2

```{r}
unique(titanic$Embarked)
```

There are 3 unique ports of embarkation: Q, S, C (Queenstown, Southampton, Cherbourg). 

## Chapter 2 {#sec-ex02-sol}

@exr-ch02-c01 b. geom_line(), c. geom_col(), e. geom_histogram()

@exr-ch02-c03 b. changing the transparency, e. jittering the points

@exr-ch02-c05 b. geom_col()

@exr-ch02-c07 b. linegraph

@exr-ch02-c09 d. boxplot

@exr-ch02-c11 The histogram is unimodal and left skewed.

@exr-ch02-app1

```{r}
#| include: false
library(ggplot2)
library(dplyr)
library(ISDSdatasets)

covid_sub <- covid_states %>% 
  filter(
    state_abbr == "IL" | state_abbr == "FL",
    date >= "2021-07-01", date <= "2021-08-31",
    wday != "Sat", wday != "Sun"
    )
```

```{r}
ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line()
```

Florida has significantly more covid cases than Illinois in the months of July and August 2021. The spikes to 0 could indicate testing did not occur on those days. Illinois has a positive "fairly linear" trend. Florida has a nonlinear trend where cases are increasing until around August 15th, at which point cases start to decrease.

@exr-ch02-app3

```{r}
ggplot(nba, aes(x = ft_percent, y = fg_percent)) +
  geom_jitter(alpha=0.1)
```

The relationship between the field goal percentage and free throw percentage seems to have no association. This means that whether a player made all of their free throws or only 50% of their free throws, it will not impact their shooting during the game (field goal percent). If we draw a circle around the points it is fairly horizontal (indicating no positive or negative trend) and a large oval. 

@exr-ch02-adv1

```{r}
ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Chapter 3 {#sec-ex03-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 4 {#sec-ex04-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::



## Chapter 5 {#sec-ex05-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 6 {#sec-ex06-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::