# Exercise solutions {#sec-exr-sol}

```{r}
library(ISDSdatasets)
library(tidyverse)
```

## Chapter 1 {#sec-ex01-sol}

@exr-ch01-c01 b. Quarto Document

@exr-ch01-c03 a. TRUE

@exr-ch01-c05 b. FALSE

@exr-ch01-c07 b. Data on a flight

@exr-ch01-app1

```{r}
z <- 12*31
add_on <- 12
z + add_on
```

@exr-ch01-adv1

```{r}
head(titanic)
```

The `head()` function shows the first 6 rows of the dataset. Based on this, I expect the `tail()` function to show the last 6 rows of the dataset.

@exr-ch01-adv2

```{r}
unique(titanic$Embarked)
```

There are 3 unique ports of embarkation: Q, S, C (Queenstown, Southampton, Cherbourg). 

## Chapter 2 {#sec-ex02-sol}

@exr-ch02-c01 b. geom_line(), c. geom_col(), e. geom_histogram()

@exr-ch02-c03 b. changing the transparency, e. jittering the points

@exr-ch02-c05 b. geom_col()

@exr-ch02-c07 b. linegraph

@exr-ch02-c09 d. boxplot

@exr-ch02-c11 The histogram is unimodal and left skewed.

@exr-ch02-app1

```{r}
#| include: false
library(ggplot2)
library(dplyr)
library(ISDSdatasets)

covid_sub <- covid_states %>% 
  filter(
    state_abbr == "IL" | state_abbr == "FL",
    date >= "2021-07-01", date <= "2021-08-31",
    wday != "Sat", wday != "Sun"
    )
```

```{r}
ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line()
```

Florida has significantly more covid cases than Illinois in the months of July and August 2021. The spikes to 0 could indicate testing did not occur on those days. Illinois has a positive "fairly linear" trend. Florida has a nonlinear trend where cases are increasing until around August 15th, at which point cases start to decrease.

@exr-ch02-app3

```{r}
ggplot(nba, aes(x = ft_percent, y = fg_percent)) +
  geom_jitter(alpha=0.1)
```

The relationship between the field goal percentage and free throw percentage seems to have no association. This means that whether a player made all of their free throws or only 50% of their free throws, it will not impact their shooting during the game (field goal percent). If we draw a circle around the points it is fairly horizontal (indicating no positive or negative trend) and a large oval. 

@exr-ch02-adv1

```{r}
ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Chapter 3 {#sec-ex03-sol}


@exr-ch03-c01 c)  `x %>% c() %>% b() %>% a()`

@exr-ch03-c03 c. 1 row and 4 columns

@exr-ch03-c05 a.  increase

@exr-ch03-c07 b.  $mean < median$

@exr-ch03-c09 d.

@exr-ch03-c11 a. e. (there is no variable called `passenger`)


@exr-ch03-app2

```{r}
library(ISDSdatasets)
```
```{r}
nba %>%
  group_by(player) %>%
  mutate(spread = pts_tm - pts_opp ) %>%
  ungroup() %>%
  slice_max(spread) %>% 
  select(player, pts_tm, pts_opp, spread)

# Alternate code
nba_spread <- nba %>%
  group_by(player) %>%
  mutate(spread = pts_tm - pts_opp ) %>% 
  select(player, pts_tm, pts_opp, spread)

nba_spread %>% arrange(desc(spread))
```
Kobe Bryant had the largest win spread, winning the game by 55 points.

@exr-ch03-app4

```{r}
covid_join <- covid_states %>%
 group_by(location_key, state) %>%
 summarize(total_new_confirmed=sum(new_confirmed, na.rm=TRUE),
           total_new_deceased=sum(new_deceased, na.rm=TRUE),
           total_new_recovered=sum(new_recovered, na.rm=TRUE),
           total_new_tested=sum(new_tested, na.rm=TRUE))

covid_joined <- covid_join %>%
  inner_join(covid_dem, by= "location_key")

covid_joined
```

@exr-ch03-adv1

```{r}
covid_pop <- covid_states %>% 
  filter(state_abbr %in% c("CA", "TX", "FL", "NY"))
```

@exr-ch03-adv2
```{r}
library(lubridate)
covid_confirmed <- covid_pop %>%
  filter(date<="2021-12-31", date>="2021-01-01") %>%
  mutate(week = week(date)) %>% 
  group_by(week, state) %>% 
  summarize(total_confirmed = sum(new_confirmed))

ggplot(data=covid_confirmed, 
       aes(x=week, y=total_confirmed, color=state)) +
  geom_line()
```

The linegraph shows that the states had the similar pattern overall. The peaks seem to correspond to weeks of holidays and school starting where people are gathering.

## Chapter 4 {#sec-ex04-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::



## Chapter 5 {#sec-ex05-sol}


@exr-ch05-c01 c) -0.7

@exr-ch05-c03 b) Between -1 and 0

@exr-ch05-c05 b) predictor variable & c) outcome variable & e) dependent variable

@exr-ch05-c07 d) For every increase of 1 unit in x, there is an associated increase of, on average, 3.86 units of y.

@exr-ch05-c09 b) FALSE

@exr-ch05-c11 d) No, the positive correlation does not necessarily imply causation.


@exr-ch05-app2
```{r}
model_pts <- lm(pts_tm ~ pts, data = nba)

summary(model_pts)
```

**a)** 0.272
```{r}
nba %>% 
  summarize(cor = cor(pts_tm, pts, use = "complete.obs"))
```

**b)**
```{r}
ggplot(nba, aes(x = pts, y = pts_tm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**c)**
$$\widehat{pts tm} = 92.80 + 0.345*pts$$
**d)**

**Intercept**: When the individual player scored 0 points in a game, we expect their team to score 92.80 points.

**Slope:**: For every 1 additional point the individual player scores, we expect the number of total points the team scores to increase by 0.345.

**e)** We predict the team will have a total of 110 points.
```{r}
92.80 + 0.345*50
```


@exr-ch05-adv1
```{r}
# coefficients
summary(model_pts)$coefficients
# or
model_pts$coefficients

# r.squared
summary(model_pts)$r.squared
```


## Chapter 6 {#sec-ex06-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::