# Exercise solutions {#sec-exr-sol}

```{r}
#| warning: false
#| message: false

library(ISDSdatasets)
library(tidyverse)
library(moderndive)
library(lubridate)
```

## Chapter 1 {#sec-ex01-sol}

@exr-ch01-c01 b. Quarto Document

@exr-ch01-c03 a. TRUE

@exr-ch01-c05 b. FALSE

@exr-ch01-c07 b. Data on a flight

@exr-ch01-app1

```{r}
z <- 12*31
add_on <- 12
z + add_on
```

@exr-ch01-adv1

```{r}
head(titanic)
```

The `head()` function shows the first 6 rows of the dataset. Based on this, I expect the `tail()` function to show the last 6 rows of the dataset.

@exr-ch01-adv2

```{r}
unique(titanic$Embarked)
```

There are 3 unique ports of embarkation: Q, S, C (Queenstown, Southampton, Cherbourg). 

## Chapter 2 {#sec-ex02-sol}

@exr-ch02-c01 b. geom_line(), c. geom_col(), e. geom_histogram()

@exr-ch02-c03 b. changing the transparency, e. jittering the points

@exr-ch02-c05 b. geom_col()

@exr-ch02-c07 b. linegraph

@exr-ch02-c09 d. boxplot

@exr-ch02-c11 The histogram is unimodal and left skewed.

@exr-ch02-app1

```{r}
#| include: false
library(ggplot2)
library(dplyr)
library(ISDSdatasets)

covid_sub <- covid_states %>% 
  filter(
    state_abbr == "IL" | state_abbr == "FL",
    date >= "2021-07-01", date <= "2021-08-31",
    wday != "Sat", wday != "Sun"
    )
```

```{r}
#| warning: false
#| message: false

ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line()
```

Florida has significantly more covid cases than Illinois in the months of July and August 2021. The spikes to 0 could indicate testing did not occur on those days. Illinois has a positive "fairly linear" trend. Florida has a nonlinear trend where cases are increasing until around August 15th, at which point cases start to decrease.

@exr-ch02-app3

```{r}
#| warning: false
#| message: false

ggplot(nba, aes(x = ft_percent, y = fg_percent)) +
  geom_jitter(alpha=0.1)
```

The relationship between the field goal percentage and free throw percentage seems to have no association. This means that whether a player made all of their free throws or only 50% of their free throws, it will not impact their shooting during the game (field goal percent). If we draw a circle around the points it is fairly horizontal (indicating no positive or negative trend) and a large oval. 

@exr-ch02-adv1

```{r}
#| warning: false
#| message: false

ggplot(
  covid_sub, 
  aes(x = date, y = new_confirmed, color = state_abbr)
  ) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Chapter 3 {#sec-ex03-sol}


@exr-ch03-c01 c)  `x %>% c() %>% b() %>% a()`

@exr-ch03-c03 c. 1 row and 4 columns

@exr-ch03-c05 a.  increase

@exr-ch03-c07 b.  $mean < median$

@exr-ch03-c09 d.

@exr-ch03-c11 a. e. (there is no variable called `passenger`)


@exr-ch03-app2

```{r}
#| warning: false
#| message: false

nba %>%
  group_by(player) %>%
  mutate(spread = pts_tm - pts_opp ) %>%
  ungroup() %>%
  slice_max(spread) %>% 
  select(player, pts_tm, pts_opp, spread)

# Alternate code
nba_spread <- nba %>%
  group_by(player) %>%
  mutate(spread = pts_tm - pts_opp ) %>% 
  select(player, pts_tm, pts_opp, spread)

nba_spread %>% arrange(desc(spread))
```
Kobe Bryant had the largest win spread, winning the game by 55 points.

@exr-ch03-app4

```{r}
#| warning: false
#| message: false

covid_join <- covid_states %>%
 group_by(location_key, state) %>%
 summarize(total_new_confirmed=sum(new_confirmed, na.rm=TRUE),
           total_new_deceased=sum(new_deceased, na.rm=TRUE),
           total_new_recovered=sum(new_recovered, na.rm=TRUE),
           total_new_tested=sum(new_tested, na.rm=TRUE))

covid_joined <- covid_join %>%
  inner_join(covid_dem, by= "location_key")

covid_joined
```

@exr-ch03-adv1

```{r}
#| warning: false
#| message: false

covid_pop <- covid_states %>% 
  filter(state_abbr %in% c("CA", "TX", "FL", "NY"))
```

@exr-ch03-adv2
```{r}
#| warning: false
#| message: false

covid_confirmed <- covid_pop %>%
  filter(date<="2021-12-31", date>="2021-01-01") %>%
  mutate(week = week(date)) %>% 
  group_by(week, state) %>% 
  summarize(total_confirmed = sum(new_confirmed))

ggplot(data=covid_confirmed, 
       aes(x=week, y=total_confirmed, color=state)) +
  geom_line()
```

The linegraph shows that the states had the similar pattern overall. The peaks seem to correspond to weeks of holidays and school starting where people are gathering.

## Chapter 4 {#sec-ex04-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::



## Chapter 5 {#sec-ex05-sol}


@exr-ch05-c01 c) -0.7

@exr-ch05-c03 b) Between -1 and 0

@exr-ch05-c05 b) predictor variable & c) outcome variable & e) dependent variable

@exr-ch05-c07 d) For every increase of 1 unit in x, there is an associated increase of, on average, 3.86 units of y.

@exr-ch05-c09 b) FALSE

@exr-ch05-c11 d) No, the positive correlation does not necessarily imply causation.


@exr-ch05-app2
```{r}
#| warning: false
#| message: false

model_pts <- lm(pts_tm ~ pts, data = nba)

summary(model_pts)
```

**a)** 0.272
```{r}
#| warning: false
#| message: false

nba %>% 
  summarize(cor = cor(pts_tm, pts, use = "complete.obs"))
```

**b)**
```{r}
#| warning: false
#| message: false

ggplot(nba, aes(x = pts, y = pts_tm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**c)**
$$\widehat{pts tm} = 92.80 + 0.345*pts$$
**d)**

**Intercept**: When the individual player scored 0 points in a game, we expect their team to score 92.80 points.

**Slope:**: For every 1 additional point the individual player scores, we expect the number of total points the team scores to increase by 0.345.

**e)** We predict the team will have a total of 110 points.
```{r}
92.80 + 0.345*50
```


@exr-ch05-adv1
```{r}
#| warning: false
#| message: false

# coefficients
summary(model_pts)$coefficients
# or
model_pts$coefficients

# r.squared
summary(model_pts)$r.squared
```


## Chapter 6 {#sec-ex06-sol}


@exr-ch06-c01 c. The parallel slopes model. Since two models are very similar, the additional complexity of the interaction model isn’t necessary

@exr-ch06-c03 b. False

@exr-ch06-c05 a. Splitting up your data can result in unequal balance in representation of some groups compared to others. & d. Splitting up your data by a confounding variable can allow you to see trends in the data that were hidden in the aggregated version of the data.


@exr-ch06-app2

```{r}
#| warning: false
#| message: false

covid_merge <- covid_states %>% 
  left_join(covid_dem)

covid_pred2 <- lm(new_confirmed ~ population + new_tested, 
                  data = covid_merge)

summary(covid_pred2)$coefficients
```



@exr-ch06-app4

```{r}
#| include: false
model_pts_parallel <- lm(pts_tm ~ pts + location, 
                   data = nba)

summary(model_pts_parallel)$coefficients

ggplot(nba, aes(x = pts, y = pts_tm, color = location)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE)
```


```{r}
#| warning: false
#| message: false

model_pts_int <- lm(pts_tm ~ pts*location, 
                   data = nba)

summary(model_pts_int)$coefficients

ggplot(nba, aes(x = pts, y = pts_tm, color = location)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

**b_0** When the star player scores 0 points and the game is away, the team is predicted to score 88.57 points.

**b_1** For every additional point the star player scores, the associated expected increase in team points is 0.431 for away games.

**b_2** When the star player scores 0 points, the team is expected to score on average 8.37 more points when the game is home compared to when the game is away.

**b_3** For every additional point the star player scores, the associated expected increase in team points will be 0.169 less for home games than away games.

@exr-ch06-adv1

```{r}
#| warning: false
#| message: false

# parallel slopes
mean(model_pts_parallel$residuals^2)

# interaction
mean(model_pts_int$residuals^2)
```
The MSE of the parallel slopes is 136.6 and the MSE of the interaction is 135.9. Since the MSE for the interaction model is smaller it appears to be slightly better.

@exr-ch06-adv2

**Many different possible answers**

```{r}
#| warning: false
#| message: false

model_better <- lm(pts_tm ~ pts_opp + pts, data = nba)

mean(model_better$residuals^2)
```
This model which uses the opponents points and star player points has an MSE of 116.9 which is better.

## Chapter 7 {#sec-ex07-sol}


@exr-ch07-c01 a) There is a non-zero probability of being selected into the treatment or control group for every unit & c) A random process is used for selection & e) A random process is used for administration of the treatments

@exr-ch07-c03 b) False

@exr-ch07-c05 a) Try to ensure that treatment and control groups are as similar as possible on all variables related to treatment assignment & c) Look for variables you can use to control for confounding & d) State your assumptions and limitations

@exr-ch07-c07 b) video games are only correlated with violent behavior, because this would be an observational study

@exr-ch07-c09 d) No, because the treatment and control groups were not randomized

@exr-ch07-app1 No the administration cannot conclude the after-school program caused student improvement. While this was a randomized selection of a subset of students we cannot generalize to all students because all students were not considered. Also, this was a before and after study where it is likely the material from the fall semester is likely different from the material in the spring semester. Perhaps these students were just better at the topics covered in the spring.


## Chapter 8 {#sec-ex08-sol}

@exr-ch08-c01 d) a population parameter

@exr-ch08-c03 all of them (a, b, c, d)

@exr-ch08-c05 d) Systematic sampling

@exr-ch08-c07 c) Cluster sampling (with unequal probability) choosing towns is based on random cluster selection

@exr-ch08-c09 c) An observational study with random sampling but no random assignment


@exr-ch08-app1
population: US citizens

parameter: proportion (most likely a Yes or No question)

undercoverage: citizens that do not own a house, if there are multiple citizens in one household only one person will receive the survey.


## Chapter 9 {#sec-ex09-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 10 {#sec-ex10-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 11 {#sec-ex11-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 12 {#sec-ex12-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::