# Exercise solutions {#sec-exr-sol}

```{r}
library(ISDSdatasets)
```

## Chapter 1 {#sec-ex01-sol}

@exr-ch01-c01 b. Quarto Document

@exr-ch01-c03 a. TRUE

@exr-ch01-c05 b. FALSE

@exr-ch01-c07 b. Data on a flight

@exr-ch01-app1

```{r}
z <- 12*31
add_on <- 12
z + add_on
```

@exr-ch01-adv1

```{r}
head(titanic)
```

The `head()` function shows the first 6 rows of the dataset. Based on this, I expect the `tail()` function to show the last 6 rows of the dataset.

@exr-ch01-adv2

```{r}
unique(titanic$Embarked)
```

There are 3 unique ports of embarkation: Q, S, C (Queenstown, Southampton, Cherbourg). 

## Chapter 2 {#sec-ex02-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 3 {#sec-ex03-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 4 {#sec-ex04-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::



## Chapter 5 {#sec-ex05-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::


## Chapter 6 {#sec-ex06-sol}

::: callout-caution
## Under Construction

Currently working on exercise solutions.
:::